String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-06 2023-1-19/sections/implementation.tex
benchmark

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
This section presents the empirical evaluation of OstrichCEA, which is our implementation of the decision procedure introduced in Section \ref{sec:algorithm}. Our objective is to validate the effectiveness of the proposed techniques by evaluating our tool's correctness and efficiency compared to other solvers. Furthermore, we assess the efficacy of our heuristics by testing different configurations of the tool. We have implemented our encoding for counting with two heuristic algorithms on Ostrich+ \cite{atva2020}. The pre-image computation for concatenation, \verb|indexOf|, \verb|substring|, \verb|replaceAll|, \verb|reverse| and finite transducer remain unchanged. OstrichCEA is written in Scala and based on the SMT solver Princess\cite{princess}.
\subsection{Benchmarks}
We conducted a comparison on four sets of benchmarks based on regex with 
counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of counting operator (the sum of upper bounds is greater than 50), and "small" contains remained instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\subsubsection{StackOverflow} is the 37381 instances we generated similarly to the \textbf{ReDos} benchmark. As the Regexlib benchmark, the real-world regex expressions are collected by James C. Davis et al.\cite{regex_lingua_franca} from StackOverflow website\cite{stackoverflow}. The website is a question-and-answer site for professional and enthusiast programmers. We extract 37,381 instances containing bounded repetition from almost 500,000 instances. The length lower bound is set to $20$ for the "small" set and $50$ for the "large" set.
\begin{equation} \label{eq:template}
  x\in \regex \wedge x\not\in \Sigma^*(<\mid >\mid '\mid ''\mid \&)\Sigma^*\wedge |x| > 50(or \ 20)
\end{equation}
\subsection{Experimental Setup and Compared Solvers}

@@

\subsection{Overall Evaluation}
In Fig.\ref{fig:cactus_all}, the cactus plot illustrates the cumulative time each solver takes for all cases in ascending order of runtime. Solvers located towards the right and lower portion of the plot indicate better performance. \newline
Table \ref{tab:results_all} summarizes the results that demonstrate OstrichCEA's superior performance, solving the most significant number of instances and outperforming most competing solvers. Including timeouts, OstrichCEA is \textbf{1.52}\mult{} faster than Z3Seq, \textbf{2.23}\mult{} faster than Ostrich, \textbf{2.26}\mult{} faster than Z3-Trau, \textbf{3.08}\mult{} faster than Z3str3, \textbf{3.11}\mult{} faster than CVC5 and close to Z3str3RE with \%2 speed loss . Note that CVC5\cite{cvc5} yielded 5370 timeouts(11\% of all instances), and Z3str3\cite{z3str3} yielded 6139 timeouts(12\% of all instances), which is much more than other solvers. Both CVC5 and Z3str3 are DPLL(T)-based solvers. It seems that almost 10\% of the benchmarks we used seem unsuitable for them. Z3-trau\cite{z3trau} yielded 21,152 unknowns because it does not support \verb|re.diff|. Z3-trau\cite{z3trau} yielded 6673 crashes (13\% of the instances) and 1233 soundness errors (2\% of the instances), which is a large portion. Z3-based solvers Z3str3 yielded 38 soundness errors, Z3seq\cite{z3seq} yielded 51 soundness errors and Z3str3RE\cite{z3str3re} yielded 39 soundness error. Most of them are due to the mistake formalization of the backslash character. OstrichCEA resulted 5 unknowns because it reached the maximum threshold of limited memory 2GB. More details of the results of each benchmark are shown in \ref{appendix:experiential_results}.
\begin{figure}[h]
  \centering
  \import{figures}{cactus_plot_all.tex}
  \caption{Cactus plot summarizing performance on all benchmarks.}
  \label{fig:cactus_all}
\end{figure}
\begin{table}
  \import{tables}{table_all.tex}
  \caption{Total results of string solvers on all benchmarks. OstrichCEA solved the most benchmarks in the second shortest time.}
  \label{tab:results_all}
\end{table}
\subsection{Analysis of Individual Heuristics}
In order to demonstrate the efficacy of the individual heuristics outlined in Section \ref{sec:algorithm} and incorporated into OstrichCEA, we assessed various tool configurations in which one or more heuristics were disabled. Figure \ref{fig:cactus_heuristics} and Table \ref{tab:results_heuristics} display the outcomes. The "OstrichCEA" plot line represents the tool's performance when all heuristics are enabled, while the "All heuristics off" line represents performance when all heuristics are disabled. The remaining plot lines exhibit performance with only the named heuristic disabled while all others are enabled. From the plots and table, it is evident that OstrichCEA functions most effectively when all heuristics are enabled. In average time with timeout times, OstrichCEA performs \textbf{1.78}$\times$ faster than employing none of our heuristics, \textbf{1.37}$\times$ faster than turning off simplification, \textbf{1.15}$\times$ faster than truing off under-approximation.
\begin{figure}
  \subimport{figures/}{cactus_plot_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics using a cactus plot.}
  \label{fig:cactus_heuristics}
\end{figure}
\begin{table}
  \subimport{tables}{table_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics.}
  \label{tab:results_heuristics}
\end{table}

>>>>

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}

We implement the algorithms in Section~\ref{sec:algorithm} on top of OSTRICH \cite{ostrich,atva2020}. Then we do experiments on two benchmark suites, to evaluate the efficacy of our approach.  In the sequel, we first describe the two benchmark suites, then present the settings and results of the experiments.

%This section presents the empirical evaluation of OstrichCEA, which is our implementation of the decision procedure introduced in Section \ref{sec:algorithm}. Our objective is to validate the effectiveness of the proposed techniques by evaluating our tool's correctness and efficiency compared to other solvers. Furthermore, we assess the efficacy of our heuristics by testing different configurations of the tool. We have implemented our encoding for counting with two heuristic algorithms on Ostrich+ \cite{atva2020}. The pre-image computation for concatenation, \verb|indexOf|, \verb|substring|, \verb|replaceAll|, \verb|reverse| and finite transducer remain unchanged. OstrichCEA is written in Scala and based on the SMT solver Princess\cite{princess}.

\subsection{Benchmark suites}

Our experiments utilize two benchmark suites, namely, AutomatArk and RegCoLe. In total, there are 49,379 instances in the two benchmark suites.

\paragraph*{AutomatArk benchmark suite.}
The AutomatArk benchmark suite is adapted from the AutomatArk suite in \cite{z3str3re} by filtering out the string constraints that contain no occurrences of counting operators and adding the length constraints. There are 8,751 instances in the AutomatArk suite.
Note that the original AutomatArk benchmark suite in \cite{z3str3re}  includes 19,979 instances, which are conjunctions of regular membership queries and generated out of regular expressions from \cite{automatark}.

\paragraph*{RegCoLe benchmark suite.} There are 40,628 RECL constraints in the RegCoLe suite.  These constraints are generated by extracting regular expressions with counting operators from \cite{regex_lingua_franca,redos_lenka} and constructing for each regular expression $e$ a RECL constraint of the form 
$$x \in e \wedge x \in \overline{\Sigma^*(<+ >+'+''+\&)\Sigma^*} \wedge |x| > 10,$$   
where $\Sigma^*(<+ >+'+''+\&)\Sigma^*$ is a regular expression that specifies at least one occurrence of special characters $<$, $>$, $'$, $''$, or $\&$, and $\overline{e'}$ denotes the complement of a regular expression $e'$. We would like to remark that there are about 500,000 real-world regular expressions in \cite{regex_lingua_franca,redos_lenka} and regular expressions with counting operators occupy about 8\% of them.

%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
\hide{
We conducted a comparison on four sets of benchmarks based on regex with 
counting operator, consisting of a total of 49,379 instances. We analyze all 19 developed benchmarks listed in \cite{zaligvinder_2021} and find that only \textbf{AutomatArk} benchmarks have regular membership with counting operator. In total, almost 18\% of the instances we evaluated were sourced from published industrial benchmarks or other solver developers. The other instances are generated by ourselves with regular expressions from real world. Each set of the benchmark is evenly divided into "large" and "small": the "large" set contains instances with large upper bounds of counting operator (the sum of upper bounds is greater than 50), and "small" contains remained instances. About 10\% of the benchmarks are in a "large" set. More details about the benchmarks are shown below.

@@

\subsubsection{StackOverflow} is the 37381 instances we generated similarly to the \textbf{ReDos} benchmark. As the Regexlib benchmark, the real-world regex expressions are collected by James C. Davis et al.\cite{regex_lingua_franca} from StackOverflow website\cite{stackoverflow}. The website is a question-and-answer site for professional and enthusiast programmers. We extract 37,381 instances containing bounded repetition from almost 500,000 instances. The length lower bound is set to $20$ for the "small" set and $50$ for the "large" set.
\begin{equation} \label{eq:template}
  x\in \regex \wedge x\not\in \Sigma^*(<\mid >\mid '\mid ''\mid \&)\Sigma^*\wedge |x| > 50(or \ 20)
\end{equation}
}
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%
%%%%%%%%%%%%%%%%%% the benchmark suite written by Denghang %%%%%%%%%%%

\subsection{Experiments}

@@

\subsection{Overall Evaluation}
In Fig.\ref{fig:cactus_all}, the cactus plot illustrates the cumulative time each solver takes for all cases in ascending order of runtime. Solvers located towards the right and lower portion of the plot indicate better performance. \newline
Table \ref{tab:results_all} summarizes the results that demonstrate OstrichCEA's superior performance, solving the most significant number of instances and outperforming most competing solvers. Including timeouts, OstrichCEA is \textbf{1.52}\mult{} faster than Z3Seq, \textbf{2.23}\mult{} faster than Ostrich, \textbf{2.26}\mult{} faster than Z3-Trau, \textbf{3.08}\mult{} faster than Z3str3, \textbf{3.11}\mult{} faster than CVC5 and close to Z3str3RE with \%2 speed loss . Note that CVC5\cite{cvc5} yielded 5370 timeouts(11\% of all instances), and Z3str3\cite{z3str3} yielded 6139 timeouts(12\% of all instances), which is much more than other solvers. Both CVC5 and Z3str3 are DPLL(T)-based solvers. It seems that almost 10\% of the benchmarks we used seem unsuitable for them. Z3-trau\cite{z3trau} yielded 21,152 unknowns because it does not support \verb|re.diff|. Z3-trau\cite{z3trau} yielded 6673 crashes (13\% of the instances) and 1233 soundness errors (2\% of the instances), which is a large portion. Z3-based solvers Z3str3 yielded 38 soundness errors, Z3seq\cite{z3seq} yielded 51 soundness errors and Z3str3RE\cite{z3str3re} yielded 39 soundness error. Most of them are due to the mistake formalization of the backslash character. OstrichCEA resulted 5 unknowns because it reached the maximum threshold of limited memory 2GB. More details of the results of each benchmark are shown in \ref{appendix:experiential_results}.
\begin{figure}[h]
  \centering
  \import{figures}{cactus_plot_all.tex}
  \caption{Cactus plot summarizing performance on all benchmarks.}
  \label{fig:cactus_all}
\end{figure}
\begin{table}
  \import{tables}{table_all.tex}
  \caption{Total results of string solvers on all benchmarks. OstrichCEA solved the most benchmarks in the second shortest time.}
  \label{tab:results_all}
\end{table}

\subsection{Analysis of Individual Heuristics}
In order to demonstrate the efficacy of the individual heuristics outlined in Section \ref{sec:algorithm} and incorporated into OstrichCEA, we assessed various tool configurations in which one or more heuristics were disabled. Figure \ref{fig:cactus_heuristics} and Table \ref{tab:results_heuristics} display the outcomes. The "OstrichCEA" plot line represents the tool's performance when all heuristics are enabled, while the "All heuristics off" line represents performance when all heuristics are disabled. The remaining plot lines exhibit performance with only the named heuristic disabled while all others are enabled. From the plots and table, it is evident that OstrichCEA functions most effectively when all heuristics are enabled. In average time with timeout times, OstrichCEA performs \textbf{1.78}$\times$ faster than employing none of our heuristics, \textbf{1.37}$\times$ faster than turning off simplification, \textbf{1.15}$\times$ faster than truing off under-approximation.
\begin{figure}
  \subimport{figures/}{cactus_plot_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics using a cactus plot.}
  \label{fig:cactus_heuristics}
\end{figure}
\begin{table}
  \subimport{tables}{table_heuristic.tex}
  \caption{A performance comparison was made on all benchmarks by disabling individual heuristics.}
  \label{tab:results_heuristics}
\end{table}
